Title: Rodent Bait Intervention to Control the Tick Population
Authors: Bruce Chidley, Sophie Stelmach, Lauren Farrell, Thaneswary Rajanderan, Geneva Liwag

This R Markdown file simulates the growth and stabilization of a population of ticks infected with Lyme disease, and how introducing acaracide bait solutions affects this population.

```{r}
require(deSolve)

mod5=function(t, y, parms){
  #Pull state variables from y vector
  L = y[1]
  Ns = y[2]
  Ni = y[3]
  A = y[4]
  Hs = y[5]
  Hi = y[6]
  Hr = y[7]
  
  #Pull the required parameter values from the parms vector
  b_l=parms["b_l"]
  b_n=parms["b_n"]
  r=parms["r"]
  a=parms["a"]
  delta=parms["delta"]
  mu_l=parms["mu_l"]
  mu_n=parms["mu_n"]
  mu_a=parms["mu_a"]
  mu_h=parms["mu_h"]
  beta=parms["beta"]
  gamma=parms["gamma"]
  gamma_h=parms["gamma_h"]
  p_hl=parms["p_hl"]
  p_nh=parms["p_nh"]
  p_hn = parms["p_hn"]
  
  #Calculation of c value for reference
  #c=(1-(1-delta)*(exp(Ni/((1-r)*(Hs+Hi+Hr)))))
  
  #Larval ticks
  dL = beta*A*exp(-gamma*A) - b_l*L*(1-r) - mu_l*L
    
  #Nymphal ticks susceptible to Lyme disease
  dNs = b_l*L*(1-r)*(1-a)*(1- 
       (1- (1-delta)^((Ni/((1-r)*(Hs+Hi+Hr)
                          )
                       )
                     )
       )
    )*((Hs + (1-p_hl)*Hi + Hr)/(Hs+Hi+Hr)) 
  - b_n*Ns*(1-r) - mu_n*Ns
    
  #Nymphal ticks infected with Lyme disease
  dNi = b_l*L*(1-r)*(1-a)*p_hl*(Hi/(Hs+Hi+Hr)) + b_l*L*(1-r)*(1-a)*( 
       (1- (1-delta)^((Ni/((1-r)*(Hs+Hi+Hr)
                          )
                       )
                     )
       )
    ) *((Hs + (1-p_hl)*Hi + Hr)/(Hs+Hi+Hr)) - b_n*Ni*(1-r) - mu_n*Ni
    
  #Adult ticks
  dA = b_n*(Ns+Ni)*(1-r)*(1-a) - mu_a*A
  
  #Hosts susceptible to Lyme disease
  dHs = mu_h*(Hs+Hi+Hr) - b_n*p_nh*(1-r)*Ni*(Hs/(Hs+Hi+Hr)) - mu_h*Hs
    
  #Hosts infected with Lyme disease
  dHi = b_n*p_nh*(1-r)*Ni*(Hs/(Hs+Hi+Hr)) - gamma_h*Hi - mu_h*Hi
    
  #Hosts Recovered from Lyme disease
  dHr = gamma_h*Hi - mu_h*Hr
  
  #Return values
  res=c(dL, dNs, dNi, dA, dHs, dHi, dHr)
  list(res)
}

```


Parameters are set such that the population reaches equilibrium (and such that the ODE solver will run properly)
```{r}
times  = seq(0, 10000, by=1)

parms= c(
  b_l=0.5,
  b_n=0.5,
  r=0.0,
  
  #Vary this "a" parameter to simulate acaracide intervention intensity. Value represents proportion of host population saturated with acaracide. Must be in the range of [0,1)
  a=0.8,
  
  delta=0.7,
  mu_l=0.01,
  mu_n=0.002,
  mu_a=0.1,
  mu_h=0.001,
  beta=15,
  gamma=0.00005,
  gamma_h=0.1,
  p_hl=0.8,
  p_nh=0.9,
  p_hn = 0.8
)
  

start = c(L=5000, Ns=2000, Ni=1000, A=500, Hs=400, Hi=1000, Hr=100)
```

System of Ordinary Differential Equations is progressed step by step according to initial conditions
```{r}
out = ode(y = start, times = times, func = mod5, 
     parms = parms)
out=as.data.frame(out)
head(round(out, 3))

out = round(out, 2)
```

Compartment populations are shown over time
```{r}
#png("intervention_60.png",width=1000,height=600)
plot(x = out$time, y = out$L, col = "black", lwd=2, ylab = "Number of individuals", 
      xlab = "Time (days)", type = "l", xlim = c(0, 120), ylim = c(0,100000))
 lines(x = out$time, y = out$Ns, lwd=2, col = "red")
 lines(x = out$time, y = out$A, lwd=2, col = "purple")
 lines(x = out$time, y = out$Hi, lwd=2, col = "magenta")
 lines(x = out$time, y = out$Ni, lwd=2, col = "pink")
 lines(x = out$time, y = out$Hr, lwd=2, col = "green")
 lines(x = out$time, y = out$Hs, lwd=2, col = "cornflowerblue")
 
 legend(x = "topright",          # Position
       legend = c("Larvae", "Sus Nymphs", "Inf Nymphs", "Adults", "Sus Hosts"
                  , "Inf Hosts", "Rec Hosts"),  # Legend texts
       #lty = c(1, 2),           # Line types
       col = c("black","red","pink","purple","cornflowerblue","magenta",
               "green"),           # Line colors
       lwd = 2) 
 
 tail(out)
```
